---
title: "Data Analysis"
author: "Abby Wallace, Lisa Woelfl, Torrence Banks, Noah Ferguson"
date: "11/11/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are analyzing data from the Maryland Lobbying Registrations. [https://lobby-ethics.maryland.gov/]

## Load libraries

```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
```

## Load and Cleaning Data

```{r}
# Loading the dataframes and binding them together

one <- read_csv("data/registrations(1).csv") %>% clean_names()
two <- read_csv("data/registrations(2).csv") %>% clean_names()
three <- read_csv("data/registrations(3).csv") %>% clean_names()
four<- read_csv("data/registrations(4).csv") %>% clean_names()
five <- read_csv("data/registrations(5).csv")%>% clean_names()
six <- read_csv("data/registrations(6).csv")%>% clean_names()
seven <- read_csv("data/registrations(7).csv")%>% clean_names()
eight <- read_csv("data/registrations(8).csv")%>% clean_names()
nine <- read_csv("data/registrations(9).csv")%>% clean_names()

total_registrations <- rbind(one, two, three, four, five, six, seven, eight, nine)

write_csv(total_registrations, "data/total_registrations.csv")

# We cleaned the data in OpenRefine and split the date column

clean_total_registrations <- read_csv("data/clean_total_registrations.csv")
clean_total_registrations[c('start', 'end')] <- str_split_fixed(clean_total_registrations$registration_period, '-', 2)
```

## Basic explorations

```{r}
glimpse(clean_total_registrations)
```

The original dataframe had 12855 rows and 5 columns. Our cleaned dataframe has 12855 rows and 9 columns (we added cleaned columns for organization name & employer and split the date column to have a start and end date for the registration).

The data looked pretty clean in OpenRefine. We found just a handful of names to clean up.

Limitations: The dataframe is pretty basic with a small number of columns. That means we will have to cross-reference the website, if we want to know how much a lobbyist is making or how much an employer is spending on lobbying. The standard dataset doesn't even include the registration matters. For our topical questions, we have to download new datasets for every registration matter we're interested in. Another limitation is that the earliest data is from 2019, which means we can't look at long-term trends.

To answer most of our questions, we will have to do additional research. For example, we can ask the number of employers and lobbyists in the energy sector and see how that changed over the last few years. To examine how much the top firms make, we have to go back to the website.

When looking into the PG county youth curfew, we won't get clear cut answers from the data alone. Trying to find a connection between the number of lobbyists for matters of juvenile law and the implementation of the curfew will require additional research on our part.

### Question 1



* **Question text**: 
* **Analysis summary**:



### Question 4

* **Question text**: How did the number of employers and lobbyists registered for "energy" change from 2018 to 2022?

Starting with the basic analysis of lobbyist registrations categorized under the registration matter "energy", we want to explore lobbying efforts in the energy sector.

Limitations: Although the general dataset on registrations goes back to 2015, datasets including the registration matter only go back to 2019.

**Analysis summary**: As our data shows, 106 lobbyists have registered for "energy" in 106, 132 in 2020, 123 in 2021 and 2022 in 115. One could have assumed that since the climate crisis has been big news in recent years that there would have been more lobbyists, which doesn't seem to be the case. The numbers have been declining since 2020. Since 2020 was the first pandemic year, there could be a connection. Employers might have had less money to spend on lobbying.

Interestingly, while the number of lobbyists declined, the number of employers grew before 2022. In 2019 there were 66 employers, in 2020 88, in 2021 85 and in 2022 there were 78 employers in the registration matter category "energy".

The 3 firms with the most employers registered for energy lobbying since 2019 are:
- Compass Government Relations Partners, LLC (10 employers)
- Perry White Ross & Jacobson (8 employers)
- Rifkin Weiner Livingston LLC (8 employers)

We found that Exxon Mobil Corporation appears as an employer for the first time in 2022. This is already interesting, as Exxon is a huge player in the energy sector nationwide. Further research on the MSEC website shows that Exxon has appeared in different categories on the past, "Environment" and "Others". 

But, according to his report, the lobbyist received $0. In fact, all available activity reports for Exxon indicate that the company has never spent a single dollar on lobbying in Maryland. 

According to Exxon's own 2020 lobbying report, the corporation spent the most in Alaska ($183,000) for "oil and gas issues, policies, and tax legislation". In Texas, Exxon lobbied around alcoholic beverage regulation, education, insurance, labor and malpractice-health care providers. These topics are not directly related to Exxon's business. It could be interesting to look into what Exxon is trying to achieve in Texas and how its lobbying efforts on these issues relate to its business model.

Lobbying report: https://corporate.exxonmobil.com/-/media/Global/Files/policy/lobbying/ExxonMobil_2020-Lobbying-Report.pdf

```{r}
## The dataset has been cleaned in OpenRefine. We clean the names and split the date column, so we have a start and an end date for the registration.

energy_registrations <- read_csv("data/clean_energy-registrations.csv") %>% clean_names()
energy_registrations[c('start', 'end')] <- str_split_fixed(energy_registrations$registration_period, '-', 2)

```

```{r}
## Then we transform our new colums from characters to dates and create a floor date to make grouping by year possible.

energy_registrations <- energy_registrations %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
```

```{r}
## To make sure we don't count some lobbyists twice in a year (they do sometimes register multiple times for short periods in a year), we use n_distinct.

year_energy_registrations <- energy_registrations %>% 
  group_by(start_year) %>%
  summarise(count = n_distinct(clean_lobbyist_registrant))
year_energy_registrations
```

```{r}
## We repeat the code, only this time we are looking for the number of employers lobbyists registered with.

year_energy_employers <- energy_registrations %>% 
  group_by(start_year) %>%
  summarise(count = n_distinct(clean_employer))
year_energy_employers
```


```{r}
## We want to know which firms had the most employers in our timeframe (2019-2022). We grouped by firms, counted employers and arranged the data.

employer_energy <- energy_registrations %>% 
  group_by(clean_organization_firm) %>% 
  summarise(count = n_distinct(clean_employer)) %>%
  arrange(desc(count))
employer_energy
```
```{r}

## Let's see who those firms work for

top_three <- energy_registrations %>% 
  filter(clean_organization_firm == c("Compass Government Relations Partners, LLC", "Perry White Ross & Jacobson", "Rifkin Weiner Livingston LLC")) %>% 
  select(clean_organization_firm, clean_employer) %>% 
  arrange(clean_organization_firm)
top_three
```


```{r}
## Now we want find out who has registered for the first time in 2022.

new_registrants <- energy_registrations %>% 
  group_by(clean_lobbyist_registrant) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_registrants

## And who these new registrants work for

new_registrants_employers <- energy_registrations %>% 
  group_by(clean_lobbyist_registrant, clean_employer) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_registrants_employers

## Then we want to know which employers registered for "energy" for the first time in 2022

new_employers <- energy_registrations %>% 
  group_by(clean_employer) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_employers

```
### Question X

* **Question text**: Who earned the most money in the latest complete lobbying year (Nov 21 - Oct 22) and which emploeyer spent the most?

In the MSEC activity reports, lobbyists have to report the compensation they have received. The employers state their total expenses. Limitations: While the data for lobbyist registrations can be filtered by registration matter, the same isn't possibly for the activity reports.

**Analysis summary**

Matthew Celentano is by far the top earner for the latest complete lobbying year (Nov 21 - Oct 22). He was compensated over 100 million dollars by the League of Life & Health Insurers of Maryland, Inc. Celentano works for the firm Funk & Bolton, P.A.. According to his LinkedIn profile, he is also the executive director of League of Life, which is a trade association for the insurance industry. (Unsurprisingly, League of Life is the employer that spent the most money in this lobbying year. It's the only employer spending over a million. Second place goes to Constellation Energy Generation, LLC, which spent over $800,000.)

In the years before, Celentano made much less money, between $190,000 and $300,000. We couldn't find any explanations for the sharp rise in the spending of League of Life, which has been paying Celentano for lobbying at least since 2019. There aren't any relevant news stories about League of Life or Celentano. The extremely high amount spent could indicate a simple typo or a bigger story. Either way, this finding is worth investigating.

Overall, 12 lobbyists made over a million, while 174 made less than one thousand dollars, 131 of those were compensated $0.

Maryland Veterinary Medical Association	spent the least money of all employers with just over $18,000.

```{r}

## The activity reports can't be filtered by energy registrations, but we want to know which lobbyist made the most money in the reporting period November 21 - October 22 and which employer spent the most money on lobbying.

## We fix the date

activity_reports <- read_csv("data/activity_reports_nov_21_oct_22.csv") %>% clean_names()
activity_reports[c('start', 'end')] <- str_split_fixed(activity_reports$reporting_period, '-', 2)

activity_reports <- activity_reports %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
activity_reports

## And transform the relevant columns to numeric

clean_activity_reports <- activity_reports %>% 
  mutate(total_expenses = gsub(",","",total_expenses)) %>% 
  mutate(total_expenses = str_remove(total_expenses, "[$]")) %>% 
  mutate(total_expenses = as.numeric(total_expenses)) %>% 
  mutate(lobbyist_compensation = gsub(",","",lobbyist_compensation)) %>% 
  mutate(lobbyist_compensation = str_remove(lobbyist_compensation, "[$]")) %>% 
  mutate(lobbyist_compensation = as.numeric(lobbyist_compensation))
clean_activity_reports

## Now we want to see the top earners among the lobbyists

top_lobbyists <- clean_activity_reports %>% 
  group_by(lobbyist_registrant) %>% 
  summarise(lobbyist_compensation = sum(lobbyist_compensation)) %>% 
  arrange(desc(lobbyist_compensation))
top_lobbyists

## Who made over a million?

over_million <- top_lobbyists %>% 
  filter(lobbyist_compensation > 1000000)
over_million

## Who made less than $1,000?

less_thousand <- top_lobbyists %>% 
  filter(lobbyist_compensation < 1000)
less_thousand

## Who was registered but did not make any money?

zero <- top_lobbyists %>% 
  filter(lobbyist_compensation == 0)
zero

## Who does the top earner Matthew Celentano work for?

celentano <- clean_activity_reports %>% 
  filter(lobbyist_registrant == "Celentano, Matthew")
celentano

## Now we want to know which employers spend the most money

top_employers <- clean_activity_reports %>% 
  group_by(employer) %>% 
  summarise(total_expenses = sum(total_expenses)) %>% 
  arrange(desc(total_expenses))
top_employers

```
```{r}
## Let's take a closer look at Celentano's activity reports

celentano_reports <- read_csv("data/celentano_activity_reports.csv") %>% clean_names()
celentano_reports[c('start', 'end')] <- str_split_fixed(celentano_reports$reporting_period, '-', 2)

## We fix the dates

celentano_reports <- celentano_reports %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
celentano_reports

## And convert the relevant columns to numeric

clean_celentano_reports <- celentano_reports %>% 
  mutate(total_expenses = gsub(",","",total_expenses)) %>% 
  mutate(total_expenses = str_remove(total_expenses, "[$]")) %>% 
  mutate(total_expenses = as.numeric(total_expenses)) %>% 
  mutate(lobbyist_compensation = gsub(",","",lobbyist_compensation)) %>% 
  mutate(lobbyist_compensation = str_remove(lobbyist_compensation, "[$]")) %>% 
  mutate(lobbyist_compensation = as.numeric(lobbyist_compensation))
clean_celentano_reports

## We want to see how much Celentano reported each year

clean_celentano_reports %>% 
  group_by(start_year) %>% 
  summarise(sum(lobbyist_compensation))

## When did he start working for League of Life?

clean_celentano_reports %>% 
  filter(employer == "League of Life & Health Insurers of Maryland, Inc.")
  
```
  
  