---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

```{r}
head(primary_18)
summary(primary_18)
colnames(primary_18)
glimpse(primary_18)
```


## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?
A1. Lots of contributions around 100 or 50. Less at the high and far ends, like a bell curve.

```{r}
wv_winred_categories <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 26 & amount <= 50 ~ "Between $26 and $50",
        amount > 51 & amount <= 100 ~ "Between $51 and $100",
        amount > 101 & amount <= 500 ~ "Between $101 and $500",
        amount >= 501 ~ "$501 and up"
      )
  )
head(wv_winred_categories)
```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. Betty Bender


```{r}
less_than_ten_contribs <- wv_winred_categories %>% filter(amount_category == "Less than $10")

head(less_than_ten_contribs)
```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. I'm wondering why donors would repeatedly donate small amounts. Bender's donations are pretty large as a sum. Are there any benefits to donating small amounts at a time? I could probably use the data to look into the connection between employer and amount or receiver and amount

Q4. Which month saw the most contributions?
A4. April

```{r}
wv_winred_contribs %>%
  mutate(month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. 6 contributions of $2,900 - they went to Congresswoman Carol Miller (WV), Steven Pallazo (Mississippi), Ryan Zinke (Montana), David McKinley (West Virginia) - we have 2 candidates.

```{r}
max_contribs <- wv_winred_categories %>% filter(amount == 2900)

head(max_contribs)
```
