---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I Learned on Sept.19

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### Interesting packages

Lubridate - this package makes it easier to work with dates and times.

### How to make a new column that calculates percentages

```{r}
# better ordering?
primary_18 %>%
  select(office, district, name_raw, party, jurisdiction, election_day, votes) %>% 
  mutate(
  percent_election_day = (election_day/votes)*100
)  %>% arrange(desc(percent_election_day))
```

### How to standardize spellings in my data

```{r}
standardized_maryland_cities <- maryland_cities %>%
  mutate(
    upper_city = str_to_upper(city)
)
```

### How to use "case when"

```{r}
maryland_winred_categories <- maryland_winred %>%
  mutate(
    amount_category = case_when(
        amount < 100 ~ "Less than $100",
        amount >= 100 & amount <= 500 ~ "Between $100 and $500",
        amount > 500 & amount <= 2900 ~ "Between $501 and $2900",
        amount > 2900 ~ "More than $2,900"
      )
  )
head(maryland_winred_categories)
```

### How to filter data/get rid of data

```{r}
selected_prince_georges <- prince_georges %>% select(office, district, name_raw, party, votes)

head(selected_prince_georges)
```


```{r}
maryland_expenses <- maryland_expenses %>% mutate(expenditure_date=mdy(expenditure_date))

head(maryland_expenses)
```


### Summarizing

I need to use group_by and summariz. Here's an example of grouping by county and calculating counts, sum and other descriptive statistics.


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```

