---
title: "Reverse Engineering Project"
author: "Abby Wallace, Lisa Woelfl, Torrence Banks, Noah Ferguson"
date: "29/10/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Why Many Americans Don't Vote](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

The data came from polling done by the marketing research company Ipsos for FiveThirtyEight. The poll was conducted in Sept. 2020 among a sample of U.S. citizens that oversampled young, Black and Hispanic respondents. 

The voter file company Aristotle connected participants to their voter file information to learn more about their voting history. 

Respondents who were only eligible to vote in three of the elections were dropped from the data.

Ultimately, the final sample used 5,836 respondents (5,239 matched the voter file, 597 who said they rarely or never voted and were eligible for at least four elections.)

Almost always vote:  People who voted in all (or all but one) of the national elections (presidential and midterm) they were eligible to vote in since 2000.

Vote sometimes: People who voted in at least two elections, but fewer than all the elections they were eligible to vote in (or all but one).

Rarely or never vote: People who voted in no elections, or just one.


```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
# Clean required data and prepare for analysis if needed. 
nonvoters_data <- read_csv("data/nonvoters_data.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [Nonvoters were more likely to have lower incomes; to be young; to have lower levels of education; and to say they don’t belong to either political party, which are all traits that square with what we know about people less likely to engage with the political system.]

* **Analysis summary**: We were able to verify 3 of the four likely attributes named above for nonvoters. By providing individual analyses through looking at rare/never voters and the various attributes, we calculated how many people of the sample data fit into those categories and found the total percentage of each each attribute contributes to likely nonvoters. We couldn't verify "young" which is likely due to that phrase being open for interpretation (we used the standard 18-24) and the total data did not have anyone younger than 22.

```{r}
nonvoters_categories <- nonvoters_data %>%
  mutate(
    Q30 = case_when(
      Q30 >= 3 ~ "neither political party",
      Q30 == 1 ~ "Republican",
      Q30 == 2 ~ "Democrat",
      )) %>%
  mutate(
    ppage = case_when(
      ppage <= 24 ~ "22-24",
      ppage > 24 ~"older than 24") )
nonvoters_categories %>%
  summarize(income_cat, educ, Q30, voter_category, ppage)
```

```{r}
nonvoters_total <- nonvoters_categories %>%
  group_by(voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n())
nonvoters_total
```

```{r}
total_nonvoters <- 1451
```

```{r}
nonvoters_party <- nonvoters_categories %>%
  group_by(Q30,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
    mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_party
```

```{r}
nonvoters_income <- nonvoters_categories %>%
  group_by(income_cat,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_income
```

```{r}
nonvoters_age <- nonvoters_categories %>%
  group_by(ppage,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_age
```

```{r}
nonvoters_edu <- nonvoters_categories %>%
  group_by(educ,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvoters_edu
```

### Sentence 2
* **Sentence text**: Our survey found, for instance, that occasional voters were slightly more likely than frequent voters to have a long-term disability (8 percent vs. 5 percent), and nonvoters were even more likely to fall into this category (12 percent). Black and Hispanic voters are also more likely to experience hurdles, perhaps in part because there tend to be fewer polling places in their neighborhoods. About 24 percent of Black respondents said that they had to stand in line for more than an hour while voting, and Hispanic respondents were more likely to say they had trouble accessing the polling place or couldn’t get off work in time to vote.

* **Analysis summary**: We were able to confirm the data on disability, showing that the nonvoters had the largest percentage of people receiving long-term disability. We did wonder about the data, because the survey also asked if people had any chronic illnesses. Depending on how you define disability, people who have chronic illnesses and don't receive long-term disability, should count as well. In the article itself, we think the authors should have made clear how narrow disability is defined in the data.

We weren't able to replicate the second percentage, that about 24% of Black respondents said they had to stand in line for more than an hour while voting. According to our calculations, it is even higher at 27%. Also, the survey asks if "you or any members of your household" had to wait in line for more than 1 hour - the article doesn't differentiate.

```{r}
never_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))

never_disability_voters
```

```{r}
sporadic_voters_total <- 2574
always_voters_total <- 1811
```

```{r}
always_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total)*100) %>%
arrange(desc(percent))

always_disability_voters
```

```{r}
sporadic_disability_voters <- nonvoters_data %>%
  group_by(Q10_1,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sporadic_voters_total)*100) %>%
arrange(desc(percent))

sporadic_disability_voters
```

```{r}
black_voters <- nonvoters_data %>%
  group_by(race) %>%
  filter(race == "Black") %>%
  summarise(count = n())
black_voters <- 932
```

```{r}
line_voters <- nonvoters_data %>%
  group_by(Q18_8,race) %>%
  filter(race == "Black") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/black_voters)*100) %>%
arrange(desc(percent))

line_voters
```

### Sentence 3

* **Sentence text**: For one thing, occasional voters (80 percent) and nonvoters (68 percent) are less likely to believe that politicians have an impact on their lives than consistent voters (84 percent). Similarly, people who vote sometimes (78 percent) or rarely (60 percent) are much less likely than voters who vote almost all of the time (85 percent) to say that there are people in politics who look like them — which could in turn make it harder to trust or identify with the candidates running for office.

* **Analysis summary**: From the question text for Q4, we assumed that respondents would be grouped as saying politicians did have an impact on their lives if they answered "A significant impact", "Somewhat of an impact" or "Just a slight impact". This produces the wrong data. Apparently the writers grouped "Just a slight impact" in with "No impact at all" - which seems questionable. Doing that we were able to get closer to the numbers in the sentence, but still didn't get the exact same values (68 vs 71 for nonvoters, for example).

The text references at least 4 different voter categories (sometimes, rarely, almost all of the time, nonvoters). The data only gives us 3 categories (rarely/never, sporadic, always). We don't have a way to differentiate between people who vote sometimes and those who rarely in our dataset. Our results for people who almost always vote varies slightly (85 vs 86%). We cant' say why.

```{r}
impact_looklike <- nonvoters_data %>%
  mutate(Q4_1 = case_when(
      Q4_1 < 3 ~ "yes impact",
      Q4_1 > 2 ~ "no impact")) %>%
  mutate(Q6 = case_when(
      Q6 < 4 ~ "yes look",
      Q6 == 4 ~ "no look"))
```

```{r}
nonvotes_impact <- impact_looklike %>%
  group_by(Q4_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvotes_impact
```

```{r}
sporadic_impact <- impact_looklike %>%
  group_by(Q4_1,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sporadic_voters_total)*100) %>%
arrange(desc(percent))
sporadic_impact
```

```{r}
always_impact <- impact_looklike %>%
  group_by(Q4_1,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total)*100) %>%
arrange(desc(percent))
always_impact
```

```{r}
nonvotes_looklike <- impact_looklike %>%
  group_by(Q6,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
arrange(desc(percent))
nonvotes_looklike
```

```{r}
always_looklike <- impact_looklike %>%
  group_by(Q6,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total)*100) %>%
arrange(desc(percent))
always_looklike
```

### Sentence 4

* **Sentence text**: This year, some of these occasional voters are also dealing with pandemic-specific challenges that could make voting less of a priority. For instance, people who vote more irregularly are a little more likely than those who almost always vote to have lost a job in the past year (13 percent vs. 10 percent) or to have worried about expenses (21 percent vs. 16 percent), although those who almost never vote are even likelier than those who sometimes vote to have reported these experiences.

* **Analysis summary**: We calculated the percentage of voters in each category that have lost their job in the past year and couldn't exactly replicate the data in the sentence. According to our calculation, 12.8% of people who vote sporadically have lost their job in the past year - which rounds up to the percentage in the sentence. It is confusing, however, that the voter group is called "people who vote more irregularly". The terms change throughout the article. We were able to reproduce the results for people who always vote as well.

We then calculated the percentages of people saying they had worried about expenses according to voter category and were able to reproduce the results in the sentence for those who vote sporadically (21%). Our results for those who always vote differ slightly (15.41% vs 16%).

We thought it was interesting how the sentence compares sporadic and consistent voters, but don't give specific figures for nonvoters, which is why we also calculated results for this group.

```{r}
never_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
never_votes_job
```

```{r}
sporadic_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sporadic_voters_total)*100) %>%
  arrange(desc(percent))
sporadic_votes_job
```

```{r}
always_votes_job <- nonvoters_data %>%
  group_by(Q11_1,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total)*100) %>%
  arrange(desc(percent))
always_votes_job
```

```{r}
never_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "rarely/never") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/total_nonvoters)*100) %>%
  arrange(desc(percent))
never_votes_worry
```

```{r}
sporadic_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "sporadic") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sporadic_voters_total*100)) %>%
  arrange(desc(percent))
sporadic_votes_worry
```

```{r}
always_votes_worry <- nonvoters_data %>%
  group_by(Q11_5,voter_category) %>%
  filter(voter_category == "always") %>%
  summarise(count = n()) %>%
  mutate(percent = (count/always_voters_total*100)) %>%
  arrange(desc(percent))
always_votes_worry
```

### Sentence 5

* **Sentence text**: Thompson thinks a lot of the problems on Election Day come down to logistical issues that could be fixed. Others agreed that there’s more the U.S. could be doing to encourage people to vote. Fifty-one percent of respondents thought that making Election Day a national holiday would allow more people to cast a ballot. Similar shares agreed that more information about the candidates from unbiased sources (51 percent) or early in-person voting (52 percent) would help, too.

* **Analysis summary**: In sentence 5, we assumed that "respondents" in this case meant the total number of people that responded to the survey without including any additional information on their race or voting status. The number of people who answered the survey was stored in the variable titled Total_Respondents. Ultimately, we were not able to get the exact results from the paragraph, as two of them were slightly off. For making Election Day a holiday, we got 52% instead of 51% and for early in-person voting, we got 53.6% instead of 52%. We were able to get 51.7% for more information on unbiased source, close to the 51% mentioned in the story. We are not sure why these numbers are slightly off.

```{r}
Total_Respondents <- 5836
```

```{r}
Holidays <- nonvoters_data %>% 
  group_by(Q19_3) %>% 
  filter(Q19_3 == "1") %>% 
  summarise(count = n()) %>% 
  mutate(percent = (count/Total_Respondents)*100)
Holidays
```

```{r}
Unbiased <- nonvoters_data %>% 
  group_by(Q19_2) %>% 
  filter(Q19_2 == "1") %>% 
  summarise(count = n()) %>% 
  mutate(percent = (count/Total_Respondents)*100)
Unbiased
```

```{r}
Early_inperson <- nonvoters_data %>% 
  group_by(Q19_6) %>% 
  filter(Q19_6 == "1") %>% 
  summarise(count = n()) %>% 
  mutate(percent = (count/Total_Respondents)*100)
Early_inperson
```
