---
title: "Texas State Profile"
author: "Lisa Wölfl"
date: "26/11/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, I'll examine donations and spending data for the two Texas party committees. I want to understand how state parties got their money and how they spent it. I'll look for possible stories around my findings.

## Load libraries

Loading required libraries for this analysis.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
```

## Load and Cleaning Data
.

```{r}
# Load required data

rep_contribs <- read_csv("data/rep_contribs.csv") %>% clean_names()
rep_disbursements <- read_csv("data/rep_disbursements.csv") %>% clean_names()

write_csv(rep_disbursements, "data/rep_disbursements.csv")

dem_contribs <- read_csv("data/dem_contribs.csv") %>% clean_names()
dem_disbursements <- read_csv("data/dem_disbursements.csv") %>% clean_names() 

write_csv(dem_disbursements, "data/dem_disbursements.csv")

clean_rep_disburs <- read_csv("data/clean_rep_disbursements.csv")
clean_dem_disburs <- read_csv("data/clean_dem_disbursements.csv")
```


### Who are the parties' donors?

How can you describe (broadly) the party’s donors? Where do the parties get their money and from what kinds of donors? Do they raise a lot in small amounts or from out of state donors?

```{r}

## Analysis of receipt data: Texas Democratic Party

total_dem_contribs <- dem_contribs %>% 
  group_by(contributor_name) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

state_dem_contribs <- dem_contribs %>% 
  group_by(contributor_state) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))
  
industry_dem_contribs <- dem_contribs %>% 
  group_by(contributor_occupation) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

count_dem_contribs <- dem_contribs %>% 
  group_by(contributor_name) %>% 
  summarise(count_donations = n()) %>% 
  arrange(desc(count_donations))

type_dem_contribs <- dem_contribs %>% 
  group_by(receipt_type_desc) %>% 
  summarise(total = sum(contribution_receipt_amount))
```

```{r}

## Analysis of receipt data: Texas Republican Party

total_rep_contribs <- rep_contribs %>% 
  group_by(contributor_name) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

state_rep_contribs <- rep_contribs %>% 
  group_by(contributor_state) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))
  
industry_rep_contribs <- rep_contribs %>% 
  group_by(contributor_occupation) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

count_rep_contribs <- rep_contribs %>% 
  group_by(contributor_name) %>% 
  summarise(count_donations = n()) %>% 
  arrange(desc(count_donations))

type_rep_contribs <- rep_contribs %>% 
  group_by(receipt_type_desc) %>% 
  summarise(total = sum(contribution_receipt_amount))
```

## What do the parties spend their money on?

```{r}
clean_dem_disburs
```



## What are the differences and similarities when comparing the parties?

## What about either donors or spending prompts further questions?