---
title: "Texas State Profile"
author: "Lisa Wölfl"
date: "26/11/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, I'll examine donations and spending data for the two Texas party committees. I want to understand how state parties got their money and how they spent it. I'll look for possible stories around my findings.

## Load libraries

Loading required libraries for this analysis.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
```

## Load and Cleaning Data
.

```{r}
# Load required data

rep_receipts <- read_csv("data/rep_receipts.csv") %>% clean_names()
rep_disbursements <- read_csv("data/rep_disbursements.csv") %>% clean_names()

write_csv(rep_disbursements, "data/rep_disbursements.csv")

dem_receipts <- read_csv("data/dem_receipts.csv") %>% clean_names()
dem_disbursements <- read_csv("data/dem_disbursements.csv") %>% clean_names() 

write_csv(dem_disbursements, "data/dem_disbursements.csv")
```


### Who are the parties' donors?

How can you describe (broadly) the party’s donors? Where do the parties get their money and from what kinds of donors? Do they raise a lot in small amounts or from out of state donors?

```{r}

## Analysis of receipt data: Texas Democratic Party

total_dem_receipts <- dem_receipts %>% 
  group_by(contributor_name) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

state_dem_receipts <- dem_receipts %>% 
  group_by(contributor_state) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))
  
industry_dem_receipts <- dem_receipts %>% 
  group_by(contributor_occupation) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

count_dem_receipts <- dem_receipts %>% 
  group_by(contributor_name) %>% 
  summarise(count_donations = n()) %>% 
  arrange(desc(count_donations))

type_dem_receipts <- dem_receipts %>% 
  group_by(entity_type_desc) %>% 
  summarise(total = sum(contribution_receipt_amount))
```

```{r}

## Analysis of receipt data: Texas Republican Party

total_rep_receipts <- rep_receipts %>% 
  group_by(contributor_name) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

state_rep_receipts <- rep_receipts %>% 
  group_by(contributor_state) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))
  
industry_rep_receipts <- rep_receipts %>% 
  group_by(contributor_occupation) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

count_rep_receipts <- rep_receipts %>% 
  group_by(contributor_name) %>% 
  summarise(count_donations = n()) %>% 
  arrange(desc(count_donations))

type_rep_receipts <- rep_receipts %>% 
  group_by(entity_type_desc) %>% 
  summarise(total = sum(contribution_receipt_amount))
```

## What do the parties spend their money on?

You may need OpenRefine to help you standardize the purpose of spending. Can you tell much about the strategy behind the spending from the data?



## What are the differences and similarities when comparing the parties?

## What about either donors or spending prompts further questions?